---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

Clear the environment to ensure that everything is defined and set the working
directory.

```{r}
rm(list=ls(all=TRUE))
setwd(paste0(COURSERADIR, "./05 - Reproducible Research/Quizes and Assigments"))
```

Download and unzip the data.

```{r}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" 
file <- "Dataset.zip"
folder <- "Activity Dataset"
if(!file.exists(file)){ download.file(url, file)} 
if(!file.exists(folder)){ unzip(file, list = FALSE, overwrite = TRUE)} 
```

Load the Activity data and the first lines and summaries the dataset.

```{r}
data <- read.csv("activity.csv")
head(data)
str(data)
```


## What is mean total number of steps taken per day?

We first sum how many steps are done in each day.

```{r}
StepsPerDay <- tapply(data$steps,data$date,sum,na.rm=TRUE)
```

Then, we can plot the histogram using the `ggplot2` library to make fancier plots.

```{r}
library(ggplot2)
qplot(StepsPerDay, bins=10, asp = 0.5)+
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

And calculate the mean and median of the variable `StepsPerDay`.

```{r}
MyMean <- format(round(mean(StepsPerDay, na.rm = TRUE), 2), nsmall = 2)
MyMedian <- median(StepsPerDay, na.rm = TRUE)
```

The mean is `r MyMean` and the median is `r MyMedian`.


## What is the average daily activity pattern?

We create a new dataframe with the daily activity pattern.

```{r}
MyAverage <- tapply(data$steps, data$interval, mean, na.rm=TRUE, simplify=T)
MyAverage_df <- data.frame(interval=as.integer(names(MyAverage)), avg=MyAverage)
```

And the time series plot.

```{r}
with(MyAverage_df,
     plot(interval, avg, type="l",col = "red",
          xlab="5min intervals",
          ylab="Average daily activity"))
```

Next is to check which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:

```{r}
MyAverage_df[which.max(MyAverage_df$avg),]

```



## Imputing missing values

The total number of missing values in the dataset is `r sum(is.na(data$steps))`.

```{r}
sum(is.na(data$steps))
```

We create a new data frame df_impute that is equal to the original dataset but with the missing data filled in (using mean for that interval for imputation):

```{r}
NAsdata <- data
NAs <- is.na(NAsdata$steps)
AvgInterval <- tapply(data$steps, data$interval, mean, na.rm=TRUE, simplify=T)
NAsdata$steps[NAs] <- AvgInterval[as.character(NAsdata$interval[NAs])]
```

Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.

```{r}
StepsPerDay2 <- tapply(NAsdata$steps, NAsdata$date, sum, na.rm=TRUE, simplify=T)

qplot(StepsPerDay2, bins=10, asp = 0.5)+
  labs(title = "Histogram of Steps per day (with missing data imputed)", x = "Steps per day", y = "Frequency")
```

And calculate the mean and median of the variable `StepsPerDay2`.

```{r}
MyMean2 <- format(round(mean(StepsPerDay2, na.rm = TRUE), 2), nsmall = 2)
MyMedian2 <- median(StepsPerDay2, na.rm = TRUE)
```

The mean is `r MyMean2` and the median is `r MyMedian2`.

## Are there differences in activity patterns between weekdays and weekends?

First we need to transform the data type of date column to date. We can use the
`lubridade' library.

```{r}
library(lubridate)
data$date <- ymd(data$date)
```
 
Choose if weekday or weekend. We can use the `dplyr' library to mutate values.

```{r}
library(dplyr)

data2<-data

data2 <- mutate(data2, weektype = ifelse(weekdays(data2$date) == "Saturday" | weekdays(data2$date) == "Sunday", "weekend", "weekday"))

data2$weektype <- as.factor(data2$weektype)

head(data2)
```

Finally make a panel plot.

```{r}

wk_df <- aggregate(steps ~ weektype+interval, data=data2, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(weektype),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_df)

```
